---
title: "Lab 9 - Multiple Linear Regression"
author: "Sara Orofino"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use this template to follow along in Lab Week 9. Each code chunk you'll need is already created and named. 

**Lab 9 Objectives:**

- Explore multivariate data (SLO housing prices)
- Perform multiple linear regression
- Assess diagnostics
- Compare different models by AIC
- Explain model outputs
- Make a nice table of regression results
- Make predictions using a final model
- Visualize predictions with original data

###1. Load packages

- tidyverse
- stargazer

```{r packages, include = FALSE}

# a. Load packages 'tidyverse' and 'stargazer':

library(tidyverse)
library(stargazer)

```

###2. Load data (slo_homes.csv) as a df called 'homes', then filter to create a data frame called 'homes_sub' that only include homes in SLO, Arroyo Grande, Santa Maria-Orcutt, and Atascadero

```{r get_data, include = FALSE}

# a. Read in data as 'homes':

homes <- read_csv("slo_homes.csv")

# b. Filter to only include cities "San Luis Obispo", "Santa Maria-Orcutt", "Atascadero", and "Arroyo Grande", and call this new subset 'homes_sub':

homes_sub <- homes %>% 
  filter(City == "Arroyo Grande" | City == "San Luis Obispo" | City == "Atascadero" |City == "Santa Maria-Orcutt")

```

###3. Go exploring (visual) + think critically about variables

*Note: It's OK to LOOK at things separately, even if you're including all in a model together!*

Example: if we want to compare distribution of housing prices by CITY (ignoring all other variables), we can do that:

```{r by_city}

# a. Calculate mean price by city
mean_by_city <- homes_sub %>% 
  group_by(City) %>% 
  summarize(
    mean = mean(Price)
  )

# b. Visualize prices by city
by_city <- ggplot(homes_sub, aes(x = Price)) +
  geom_density(aes(color = City, fill = City), alpha = 0.3) + # Note: just to show what the geom_violin shows
  theme_classic() +
  scale_x_continuous(expand = c(0,0), limits = c(0,3e6)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Home Prices (USD)", y = "Density")

by_city

```

Or another question: Overall relationship between home square footage and price, separated by City? 

```{r by_sqft}

# a. Relationship between square footage and price
by_sqft <- ggplot(homes_sub, aes(x = SqFt, y = Price)) +
  geom_point(aes(color = City, pch = Status), alpha = 0.5) +
  facet_wrap(~Status)

by_sqft

# Observations here: Does relationship appear ~ linear? Anything else we can pick out re: trends, outliers, etc.? What is the general trend? Any outliers? Is there reason enough for us to omit it?


# Linear relationship appears to exist between continuous variable input (square foot) and continuous variable output (housing price)

```

###4. Multiple linear regression

Multiple linear regression in R follows the same syntax we've been using so far: 

    lm(y ~ x1 + x2 + x3..., data = df_name)
    
Goal is to describe the price of a house based on a series of predictor variables.
Let's try this model a couple of different ways: 

(1) Use all available variables (saturated model) 
(2) Use only SqFt as a predictor for "home size" generally (omit Bathrooms and Bedrooms), and omit PricePerSqFt (since it's derived from two other existing variables in the model)

Use summary() to view the model output and statistics.

```{r saturated}

# a. Saturated model: 

homes_lm1 <- lm(Price ~ City + Bedrooms + Bathrooms + SqFt + PricePerSqFt + Status, data = homes_sub)

# THIS IS BAD - Price per Sqft is a variable that is calculated from the dependent variable price and this can lead to biased overfitting of model results.

# b. Model summary:

summary(homes_lm1)

# Reference levels: Status - Foreclosure; City - Arroyo Grande  
# Colinearity - if multiple variables exist in a model and they are highly correlated then including all of them doesn't make sense. (Ex - bedrooms/bathrooms/square-feet)
# Colinearity can really mess with your coefficients so if you're model doesn't make that much sense look for these. 
```

The next model: Exclude price per square foot, and bedrooms/bathrooms (since these are largely explained by square footage...)

```{r subset}

# a. Updated model with variables City, SqFt, and Status:

homes_lm2 <- lm(Price ~ City + SqFt + Status, data = homes_sub)

# b. Model summary:

summary(homes_lm2)

# p-values for each coefficient are telling us if that variable is a significant predictor of housing price --> Exs: StatusShortSale = 0.9447 (not a significant predictor), StatusRegular <0.001 (is a significant predictor)

# Adjusted R2 = 0.5268 (use when doing multiple linear regression) 

# model p-value < 0.001 --> says this model significantly predicts housing prices 

```
*Interpretation Examples:*  
All else being equal, a house in SLO would be expected to sell, on average, for $34525.40 more than a house in Arroyo Grande.   

For every 1 sqft increase in housing size, we would expect the price of a house to go up by $245.90, if all other variables are equal.  

Wait...but what if I wanted everything to be with respect to a Regular sale status? Then I need to change my factor levels. We've done this before, here we'll use a different function (fct_relevel) from *forcats* package in the tidyverse. 

```{r fct_relevel}

# a. Set Status to class 'factor'

homes_sub$Status <- factor(homes_sub$Status)

# b. Check to ensure it's a factor now
# Use class() in the console to check that it's a factor

# c. Check levels:

levels(homes_sub$Status)

# d. Reassign reference level of "Status" to "Regular":

homes_sub$Status <- fct_relevel(homes_sub$Status, "Regular")

# Recheck the levels to ensure that the reference level was reassigned (should be listed first)

levels(homes_sub$Status)

# e. Now run the regression again - same equation, but now the reference level is different (Status = Regular): 

homes_lm3 <- lm(Price ~ City + SqFt + Status, data = homes_sub)

summary(homes_lm3)

```

Interpret the coefficients and statistical outcomes above. 
Notes: 

###5. Model diagnostics

Remember, since we're concerned about *residuals* (distance that actual observations exist from model predictions), we can only evaluate some assumptions *after* running the regression. 

Then we can evaluate model diagnostics using the plot() function:

```{r diagnostics}

# a. Model diagnostics:
plot(homes_lm3)

#Notes on model diagnostics:  

# Residuals vs. Fitted - generally evenly distributed vertically around the line; definitely are a few outliers but that isn't reason enough not to include them. (satisfied)  

#QQ Plot - Following a very normal residual distribution, generally (satisfied)

# Overall, heterodascitity is oK, residuals normality definitey looks good, and conceptually and mathematically my model makes sense!
```

###6. Model comparison by Akaike Information Criterion

The AIC is a quantitative metric for model "optimization" that balances complexity with model fit. The best models are the ones that fit the data as well as possible, as simply as possible. Recall: lower AIC value indicates a *more optimal* balance - **BUT STATISTICS IS NO SUBSTITUTE FOR JUDGEMENT!!!**

```{r AIC}

# a. AIC values for each model

# Find saturated model AIC
sat_aic <- AIC(homes_lm1)

# Find final model AIC
final_aic <- AIC(homes_lm3)

# Answer: which would you pick? 
# While the absolute AIC value is lower for model 1, I would still pick model 3 (homes_lm3) because conceptually this makes sense with what I know about the variables and I know that heterodasciticity and normality are met. 


```

###7. Regression tables with *stargazer*

```{r stargazer, results = 'asis'}

# a. Prepare a nice regression table:

#once models are created all you have to do is list the model names in stargazer()

lm_table <- stargazer(homes_lm1, homes_lm3, type = "html")

# Note: If you want to work with this in Word, save to html, open, copy and paste into Word. 

```

###8. Making predictions

Using your final selected model, predict the housing price for a range of home sizes, sale status, and city. 

The predict() function uses the following syntax:

      predict(model_name, newdata = new_data_name)
      
Defaults are to exclude the prediction SE and mean confidence interval - if you want to include, use arguments

      se.fit = TRUE
      interval = "confidence" 
      interval = "prediction"

First, you need to create a new data frame of values that contain ALL NECESSARY VARIABLES **with the same variable names AND level strings**.

```{r df_new}

# First, make a new data frame

# Note that the df_new created below has the SAME variable names and level strings as the original model data (otherwise R won't know how to use it...)
# Work through this on your own to figure out what it actually does:

df_new <- data.frame(City = rep(c("San Luis Obispo",
                                  "Santa Maria-Orcutt",
                                  "Atascadero",
                                  "Arroyo Grande"), 
                                each = 60), 
                     SqFt = rep(seq(from = 500,
                                    to = 3500, 
                                    length = 20), 
                                times = 12), 
                     Status = rep(c("Regular",
                                    "Foreclosure",
                                    "Short Sale"), 
                                  times = 12, 
                                  each = 20))

# Looks at 60 observations from each city, on sequence of sqft between 500 and 3500 by 20 ft, and each status 20 times (times = 12 to match the number of sqft predictions) 
```

Make predictions for the new data using predict():

```{r predict}

# a. Make predictions using the new data frame:
# se.fit includes standard error for mean price and interval = gives a confidence interval for the mean price

price_predict <- predict(homes_lm3, newdata = df_new, se.fit = TRUE, interval = "confidence")

# b. Bind predictions to the data to make it actually useful:
# data.frame combines by column as long as they have the same number of rows.

predict_df <- data.frame(df_new, price_predict)

```

Then visualize it!

```{r graph, echo = FALSE}

# Create a line graph with square footage on the x-axis, predicted price on y-axis, with color dependent on City and facet_wrapped by sale status (follow along):

predict_graph <- ggplot(predict_df, aes(x = SqFt, y = fit.fit)) +
  geom_line(aes(color = City)) +
  facet_wrap(~ Status)

predict_graph
```

END LAB